---
title: "individual1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)

defense <- read_excel('C:/Users/tyler/OneDrive/Desktop/cyclonesFootball2020.xlsx', sheet='Defensive')
str(defense)

offense <- read_excel('C:/Users/tyler/OneDrive/Desktop/cyclonesFootball2020.xlsx', sheet='Offensive')
str(defense)

biography <- read_excel('C:/Users/tyler/OneDrive/Desktop/cyclonesFootball2020.xlsx', sheet='Biography')
str(defense)

```

1.1
```{r}
defense$Name <- as.factor(defense$Name)
  
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)

offense$Name <- as.factor(offense$Name)
  
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)

biography$Name <- as.factor(biography$Name)

```

1.2
```{r}
library(dplyr)
defense <- defense%>%
  mutate_if(is.character, as.numeric)

offense <- offense%>%
  mutate_if(is.character, as.numeric)

biography$Weight <- as.numeric(biography$Weight)

```

1.3

```{r}
library(tidyr)
biography <- biography %>%
  separate(Height,c("Feet", "Inches"), sep="-", remove= FALSE)
```
```{r}
  biography$Feet <- as.numeric(biography$Feet)
  biography$Inches <- as.numeric(biography$Inches)
  biography$Height <- (biography$Feet *12) + biography$Inches
  
  biography1= subset(biography, select = -c(Feet,Inches) )
  
```
```{r}
bioClean <-biography1
defClean <- defense
offClean <-offense
offClean
```

2.1
```{r}
offClean1 <-offClean %>%
  pivot_longer(!c(Name,Opponent_Opponent), names_to = "stat", values_to = "count" )

offClean1
```
2.2
```{r}
library(ggplot2)

ggplot(data = offClean1, aes(x=stat, y=count)) +
  geom_boxplot() + theme(axis.text.x=element_text(angle=90, hjust=1))
```

One of the noticeable major trends is the big difference between Q3 and the max value for multiple variables this shows that sometimes there can be major runs or passes for a long distance compared to the more common short gain.
```{r}
offClean2 <- offClean1[grepl("Receiving_YDS", offClean1$stat),]
offClean2$Opponent_Opponent <- as.character(offClean2$Opponent_Opponent)
offClean2 <- offClean2[!grepl("Louisiana", offClean2$Opponent_Opponent),]
offClean2 <- offClean2[!grepl("TCU", offClean2$Opponent_Opponent),]
offClean2 <- offClean2[!grepl("Texas Tech", offClean2$Opponent_Opponent),]
offClean2 <- offClean2[!grepl("Oklahoma St.", offClean2$Opponent_Opponent),]
offClean2 <- offClean2[!grepl("Kansas", offClean2$Opponent_Opponent),]
offClean2 <- offClean2[!grepl("Texas", offClean2$Opponent_Opponent),]
offClean2 <- offClean2[!grepl("West Virginia", offClean2$Opponent_Opponent),]
offClean2 <- offClean2[!grepl("Baylor", offClean2$Opponent_Opponent),]
offClean2
ggplot(offClean2, aes(x=Opponent_Opponent, y=count)) +geom_point()
```
ISU Played better against Oklahoma, the count for high receiving yards is much higher.

2.4
```{r}
bioClean1 <- bioClean %>%
  separate(Hometown,c("City", "State"), sep=", ", remove= TRUE)

bioClean1
```
2.5
```{r}
count(bioClean1, vars=State)

```
3.1

```{r}
ggplot(data=bioClean1, aes(x=State)) +geom_bar()

```
Majority of players are from Iowa 

3.2
```{r}
off <- offClean
def <- defClean

off[is.na(off)] <- 0
def[is.na(def)] <- 0

off$Super_Score <- off$Receiving_REC + off$Receiving_YDS + off$Receiving_TD + off$Rushing_ATT + off$Rushing_YDS+ off$Rushing_TD+ off$`Passing_CMP-ATT`+ off$Passing_YDS+ off$Passing_TD+ off$Passing_INT 



def$Super_Score <- def$Tackles_Solo + def$Tackles_ASST + def$Tackles_TFL + def$Tackles_Sack + def$Turnover_FF + def$Turnover_FR + def$Turnover_INT + def$Pass_QBH + def$Pass_PB

keep <- c("Name","Super_Score")

def = def[keep]

off = off[keep]

plcomp <- rbind(def, off)

plcomp <-plcomp %>%
  group_by(Name) %>%
  summarize(Score = sum(Super_Score)) %>%
  arrange(Score)


ggplot(data = plcomp, aes(x=Name, y=Score)) +
  geom_boxplot() + theme(axis.text.x=element_text(angle=90, hjust=1))

```
Players earned points based off there stats a value of 1 in a stat earned them 1 point for example. Brock Purdy out performed Every other player significantly and is the beat the mean score by about 3000 points 



